Concurrent Programming with Harmony

### 2020-07-30 book/paper.tex
Latest

\section*{Exercises}
\begin{problems}
\item Change the code in \autoref{fig:boundedbuffer} so that
\texttt{produce()} fails if the buffer is full rather than block.
\texttt{produce()} should return \texttt{False} if it fails
and return \texttt{True} if it succeeds.
\item Implement a \emph{bounded stack}.  Method \texttt{push}(\textit{item})
should block if the stack is full.  Method \texttt{pop()} should block
until the stack is non-empty and then return an item.
Also implement test programs for your solution.
\item \label{ex:gpu} Implement a thread-safe \emph{GPU allocator} by modifying
\autoref{fig:gpu}.
There are \texttt{N} GPUs identified by the numbers
1 through \texttt{N}.  Method \texttt{gpuAlloc()} returns the identifier
of an available GPU, blocking if there is currently no available GPU.
Method \texttt{gpuRelease}(\textit{gpu}) releases the given GPU.  It never needs
to block.
\end{problems}

@@

Processes can be \emph{interrupted}.  An interrupt is a notification of some event
such as a keystroke, a timer expiring, the reception of a network packet, the completion
of a disk operation, and so on.  We distinguish \emph{interrupts} and \emph{exceptions}.
An exception is caused by the process executing an invalid machine instruction such as
divide-by-zero.  An interrupt is caused by some peripheral device and can be handled in
Harmony; an exception is an error.
\index{exception}

@@

Instead, the way one fixes interrupt-safety issues is through disabling interrupts
temporarily.  In Harmony, this can be done by setting the \textit{interrupt level}
of a process to \texttt{True} using the \textbf{setintlevel} interface.
\autoref{fig:trap4} illustrates how this is done.
Note that it is not necessary to change the interrupt level during servicing an
interrupt, because it is automatically set to \texttt{True} upon entry to the interrupt
handler and restored to \texttt{False} upon exit.
It is important that the \textit{main}() code re-enables interrupts after incrementing
\textit{count}.  What would happen if \textit{main}() left interrupt disabled?

@@

It will often be necessary to write code that is both interrupt-safe \emph{and}
thread-safe.  As you might expect, this involves both managing locks and
interrupt levels.
To increment \textit{count}, the interrupt level must be \textit{True} and
\textit{countlock} must be held.
\autoref{fig:trap6} gives an example of how this might be done.
One important rule to remember is for a process to disable interrupts \emph{before}
attempting to acquire a lock.

@@

There is another important rule to keep in mind.  Just locks should never be held
for long, interrupts should never be disabled for long.  With locks the issue is to
maximize concurrent performance.  The interrupts the issue is fast response to
asynchronous events.  Because interrupts may be disabled only briefly, interrupt
handlers must run quickly and cannot wait for other events.  So it is ok to invoke
synchronization calls such as \texttt{V} and \texttt{notify}, but calls such as
\texttt{P} and \texttt{wait} should only be used if it's known that they will not
block for long.  Informally, interrupt handlers can be
\emph{producers} but not \emph{consumers} of synchronization events.

\section*{Exercises}
\begin{problems}
\item The \texttt{produce} method in \autoref{fig:boundedbuffer} cannot be used
in interrupt handlers for two reasons: (1) it is not interrupt-safe, and (2)
it may block for a long time if the buffer is full.  Yet, it would be useful if,
say, a keyboard interrupt handler could place an event on a shared queue.  Implement
a new method \texttt{iproduce}(\textit{item}) that is interrupt-safe and does not
block.  Instead, it should return \texttt{False} if the buffer is full and \texttt{True}
if the item was successfully enqueued.
\end{problems}

>>>>

\section*{Exercises}
\begin{problems}
\item \label{ex:nbproduce} Change the code in \autoref{fig:boundedbuffer} so that
\texttt{produce()} fails if the buffer is full rather than block.
\texttt{produce()} should return \texttt{False} if it fails
and return \texttt{True} if it succeeds.
\item Implement a \emph{bounded stack}.  Method \texttt{push}(\textit{item})
should block if the stack is full.  Method \texttt{pop()} should block
until the stack is non-empty and then return an item.
Also implement test programs for your solution.
\item \label{ex:gpu} Implement a thread-safe \emph{GPU allocator} by modifying
\autoref{fig:gpu}.
There are \texttt{N} GPUs identified by the numbers
1 through \texttt{N}.  Method \texttt{gpuAlloc()} returns the identifier
of an available GPU, blocking if there is currently no available GPU.
Method \texttt{gpuRelease}(\textit{gpu}) releases the given GPU.  It never needs
to block.
\end{problems}

@@

Processes can be \emph{interrupted}.  An interrupt is a notification of some event
such as a keystroke, a timer expiring, the reception of a network packet, the completion
of a disk operation, and so on.  We distinguish \emph{interrupts} and \emph{exceptions}.
An exception is caused by the process executing an invalid machine instruction such as
divide-by-zero.  An interrupt is caused by some peripheral device and can be handled in
Harmony.
In other words: an interrupt is a notification, while an exception is an error.
\index{exception}

@@

Instead, the way one fixes interrupt-safety issues is through disabling interrupts
temporarily.  In Harmony, this can be done by setting the \textit{interrupt level}
of a process to \texttt{True} using the \textbf{setintlevel} interface.
\autoref{fig:trap4} illustrates how this is done.
Note that it is not necessary to change the interrupt level during servicing an
interrupt, because it is automatically set to \texttt{True} upon entry to the interrupt
handler and restored to \texttt{False} upon exit.
It is important that the \textit{main}() code re-enables interrupts after incrementing
\textit{count}.  What would happen if \textit{main}() left interrupts disabled?

@@

It will often be necessary to write code that is both interrupt-safe \emph{and}
thread-safe.  As you might expect, this involves both managing locks and
interrupt levels.
To increment \textit{count}, the interrupt level must be \textit{True} and
\textit{countlock} must be held.
\autoref{fig:trap6} gives an example of how this might be done.
One important rule to remember is that a process should disable interrupts \emph{before}
attempting to acquire a lock.

@@

There is another important rule to keep in mind.  Just like locks should never be held
for long, interrupts should never be disabled for long.  With locks the issue is to
maximize concurrent performance.  For interrupts the issue is fast response to
asynchronous events.  Because interrupts may be disabled only briefly, interrupt
handlers must run quickly and cannot wait for other events.  So it is ok to invoke
synchronization calls such as \texttt{V} and \texttt{notify}, but calls such as
\texttt{P} and \texttt{wait} should only be used if it is certain that they will not
block for long.  Informally, interrupt handlers can be
\emph{producers} but not \emph{consumers} of synchronization events.

\section*{Exercises}
\begin{problems}
\item The \texttt{produce} method in \autoref{fig:boundedbuffer} cannot be used
in interrupt handlers for two reasons: (1) it is not interrupt-safe, and (2)
it may block for a long time if the buffer is full.  Yet, it would be useful if,
say, a keyboard interrupt handler could place an event on a shared queue.
Similar to \autoref{ex:nbproduce}, implement
a new method \texttt{iproduce}(\textit{item}) that does not
block.  Instead, it should return \texttt{False} if the buffer is full and \texttt{True}
if the item was successfully enqueued.
The method also needs to be interrupt-safe.
\end{problems}
